---
title: "Exploratory Data Analysis on Burning Glass Data v2"
output: html_document
---

```{r}
setwd("C:/Users/chari/Desktop/CSET/AI Hype Project/v2")
rm(list=ls())
library(tidyverse)
```


What degrees are most sought after for AI jobs in each sector?
```{r}
# 2019
major_sector_19 <- read_csv("major_sector_2019.csv")

# Remove null data
major_sector_19 <- filter(major_sector_19, !is.na(major_sector_19$major))

# Group by sector and order from highest job count to lowest
major_sector_19 <- major_sector_19[order(major_sector_19$naics2_name, 
                                   major_sector_19$job_count, 
                                   decreasing = T),]

# Calculate percentage, pick the top 5 majors
major_sector_19 %>%
  group_by(naics2_name) %>%
  mutate(sum = sum(job_count), percentage = job_count/sum*100) %>%
  group_by(naics2_name) %>%
  slice(1:5) -> major_sector_19.final

# Stacked bar graph
ggplot(major_sector_19.final, aes(x = naics2_name, 
                                  y = percentage, 
                                  fill = major, 
                                  label = major)) + 
  geom_bar(stat = "identity", color="black") +
  geom_text(position = position_stack(vjust = 0.5)) +
  theme_gray(base_size = 10) +
  theme(legend.position="none") +
  labs(title = "AI Labor Demand Broken Down by Major in Each Sector (2019)",
       subtitle = "(Top 5 Majors)",
       x = "Sector",
       y = "% of Job Postings (in Each Sector)")



# Aggregate
major_sector_agg <- read_csv("major_sector_aggregate.csv")

# Remove null data
major_sector_agg <- filter(major_sector_agg, !is.na(major_sector_agg$major))

# Group by sector and order from highest job count to lowest
major_sector_agg <- major_sector_agg[order(major_sector_agg$naics2_name, 
                                     major_sector_agg$job_count, 
                                     decreasing = T),]

# Calculate percentage, pick the top 5 majors
major_sector_agg %>%
  group_by(naics2_name) %>%
  mutate(sum = sum(job_count), percentage = job_count/sum*100) %>%
  group_by(naics2_name) %>%
  slice(1:5) -> major_sector_agg.final

# Stacked bar graph
ggplot(major_sector_agg.final, aes(x = naics2_name, 
                                   y = percentage, 
                                   fill = major, 
                                   label = major)) + 
  geom_bar(stat = "identity", color="black") +
  geom_text(position = position_stack(vjust = 0.5)) +
  theme_gray(base_size = 10) +
  theme(legend.position="none") +
  labs(title = "AI Labor Demand Broken Down by Major in Each Sector (2010-2020)",
       subtitle = "(Top 5 Majors)",
       x = "Sector",
       y = "% of Job Postings (in Each Sector)")

```


What degrees are most sought after for each type of AI job?
```{r}
# 2019
major_field_19 <- read_csv("major_field_2019.csv")

# Remove null data
major_field_19 <- filter(major_field_19, !is.na(major_field_19$major))

# Group by field and order from highest job count to lowest
major_field_19 <- major_field_19[order(major_field_19$DETAIL, 
                                   major_field_19$job_count, 
                                   decreasing = T),]

# Calculate percentage and pick the top 5 majors
major_field_19 %>%
  group_by(DETAIL) %>%
  mutate(sum = sum(job_count), percentage = job_count/sum*100) %>%
  group_by(DETAIL) %>%
  slice(1:5) -> major_field_19.final

major_field_19.final$DETAIL <- factor(major_field_19.final$DETAIL,
                                    levels = c("C1", "C2", "PT", "CT"))

# Stacked bar graph
ggplot(major_field_19.final, aes(x = DETAIL, 
                           y = percentage, 
                           fill = major, 
                           label = major)) + 
  geom_bar(stat = "identity", color="black") +
  geom_text(position = position_stack(vjust = 0.5)) +
  theme_gray(base_size = 15) +
  theme(legend.position="none") +
  labs(title = "AI Labor Demand Broken Down by Major in Each AI Workforce Category (2019)",
       subtitle = "(Top 5 Majors)",
       x = "AI Workforce Category",
       y = "% of Job Postings (in Each Field)")


# Aggregate
major_field_agg <- read_csv("major_field_aggregate.csv")

# Remove null data
major_field_agg <- filter(major_field_agg, !is.na(major_field_agg$major))

# Group by sector and order from highest job count to lowest
major_field_agg <- major_field_agg[order(major_field_agg$DETAIL,
                                   major_field_agg$job_count, 
                                   decreasing = T),]

# Calculate percentage, pick the top 5 majors
major_field_agg %>%
  group_by(DETAIL) %>%
  mutate(sum = sum(job_count), percentage = job_count/sum*100) %>%
  group_by(DETAIL) %>%
  slice(1:5) -> major_field_agg.final

major_field_agg.final$DETAIL <- factor(major_field_agg.final$DETAIL,
                                    levels = c("C1", "C2", "PT", "CT"))


# Stacked bar graph
ggplot(major_field_agg.final, aes(x = DETAIL, 
                                  y = percentage, 
                                  fill = major, 
                                  label = major)) + 
  geom_bar(stat = "identity", color="black") +
  geom_text(position = position_stack(vjust = 0.5)) +
  theme_gray(base_size = 15) +
  theme(legend.position="none") +
  labs(title = "AI Labor Demand Broken Down by Major in Each AI Workforce Category (2010-2020)",
       subtitle = "(Top 5 Majors)",
       x = "AI Workforce Category",
       y = "% of Job Postings (in Each Sector)")
```


How has labor demand for AI skills changed over the years in each sector?
```{r}
trend_sector <- read_csv("trend_sector.csv")

# Line graph
ggplot(trend_sector, aes(x = year, y = job_count, color = naics2_name)) +
  geom_line() + geom_point(size = 5) +
  scale_x_continuous(breaks=c(2010,2011,2012,2013,2014,2015,2016,2017,2018,2019,2020)) +
  theme_gray(base_size = 15) +
  labs(title = "Trend in AI Labor Demand by Sector",
       x = "Year",
       y = "Job Count") +
  scale_color_discrete("Sector")
```


How has labor demand for AI skills changed over the years for each type of AI job?
```{r}
trend_field <- read_csv("trend_field.csv")

# Line graph
ggplot(trend_field, aes(x = year, y = job_count, color = DETAIL)) +
  geom_line() + geom_point(size = 5) +
  scale_x_continuous(breaks=c(2010,2011,2012,2013,2014,2015,2016,2017,2018,2019,2020)) +
  theme_gray(base_size = 15) +
  labs(title = "Trend in AI Labor Demand by AI Workforce Category",
       x = "Year",
       y = "Job Count") +
  scale_color_discrete("AI Workforce Category", breaks = c("C1","C2","PT","CT"))
```


Trend in AI labor demand over the years, analyze by sector and AI workforce category
```{r}
trend_sectorField <- read_csv("trend_sector-field.csv")
trend_sectorField_total <- read_csv("trend_sector-Field_total.csv")
# Calculate total job count for each sector in each year by adding job count from all the AI workforce categories + null (which represents all non-AI jobs)
trend_sectorField_total %>%
  arrange(naics2_name, year) %>%
  group_by(naics2_name, year) %>%
  mutate(job_count = ifelse(is.na(DETAIL), sum(job_count), job_count),
         DETAIL = ifelse(is.na(DETAIL), "Total", DETAIL)) -> trend_sectorField_total


# Line graph for Manufacturing
trend_sectorField %>%
  filter(naics2_name == "Manufacturing") %>%
  ggplot(aes(x = year, y = job_count, color = DETAIL)) +
  geom_line() + geom_point(size = 5) +
  scale_x_continuous(breaks=c(2010,2011,2012,2013,2014,2015,2016,2017,2018,2019,2020)) +
  theme_gray(base_size = 15) +
  labs(title = "Trend in AI Labor Demand by AI Workforce Category (Manufacturing)",
       x = "Year",
       y = "Job Count") +
  scale_color_discrete("AI Workforce Category", breaks = c("C1","C2","PT","CT"))

# With total trend
trend_sectorField_total %>%
  filter(naics2_name == "Manufacturing") %>%
  ggplot(aes(x = year, y = job_count, color = DETAIL)) +
  geom_line() + geom_point(size = 5) +
  scale_x_continuous(breaks=c(2010,2011,2012,2013,2014,2015,2016,2017,2018,2019,2020)) +
  theme_gray(base_size = 15) +
  labs(title = "Trend in AI Labor Demand by AI Workforce Category (Manufacturing)",
       x = "Year",
       y = "Job Count") +
  scale_color_discrete("AI Workforce Category", 
                       breaks = c("Total","C1","C2","PT","CT"))


# Line graph for Professional, Scientific, and Technical Services
trend_sectorField %>%
  filter(naics2_name == "Professional, Scientific, and Technical Services") %>%
  ggplot(aes(x = year, y = job_count, color = DETAIL)) +
  geom_line() + geom_point(size = 5) +
  scale_x_continuous(breaks=c(2010,2011,2012,2013,2014,2015,2016,2017,2018,2019,2020)) +
  theme_gray(base_size = 15) +
  labs(title = "Trend in AI Labor Demand by AI Workforce Category (Professional, Scientific, and Technical Services)",
       x = "Year",
       y = "Job Count") +
  scale_color_discrete("AI Workforce Category", breaks = c("C1","C2","PT","CT"))

# With total trend
trend_sectorField_total %>%
  filter(naics2_name == "Professional, Scientific, and Technical Services") %>%
  ggplot(aes(x = year, y = job_count, color = DETAIL)) +
  geom_line() + geom_point(size = 5) +
  scale_x_continuous(breaks=c(2010,2011,2012,2013,2014,2015,2016,2017,2018,2019,2020)) +
  theme_gray(base_size = 15) +
  labs(title = "Trend in AI Labor Demand by AI Workforce Category (Professional, Scientific, and Technical Services)",
       x = "Year",
       y = "Job Count") +
  scale_color_discrete("AI Workforce Category", 
                       breaks = c("Total","C1","C2","PT","CT")) +
  scale_y_continuous(labels = function(x) format(x, scientific = FALSE))


# Line graph for Transportation and Warehousing
trend_sectorField %>%
  filter(naics2_name == "Transportation and Warehousing") %>%
  ggplot(aes(x = year, y = job_count, color = DETAIL)) +
  geom_line() + geom_point(size = 5) +
  scale_x_continuous(breaks=c(2010,2011,2012,2013,2014,2015,2016,2017,2018,2019,2020)) +
  theme_gray(base_size = 15) +
  labs(title = "Trend in AI Labor Demand by AI Workforce Category (Transportation and Warehousing)",
       x = "Year",
       y = "Job Count") +
  scale_color_discrete("AI Workforce Category", breaks = c("C1","C2","PT","CT"))

# With total trend
trend_sectorField_total %>%
  filter(naics2_name == "Transportation and Warehousing") %>%
  ggplot(aes(x = year, y = job_count, color = DETAIL)) +
  geom_line() + geom_point(size = 5) +
  scale_x_continuous(breaks=c(2010,2011,2012,2013,2014,2015,2016,2017,2018,2019,2020)) +
  theme_gray(base_size = 15) +
  labs(title = "Trend in AI Labor Demand by AI Workforce Category (Transportation and Warehousing)",
       x = "Year",
       y = "Job Count") +
  scale_color_discrete("AI Workforce Category", 
                       breaks = c("Total","C1","C2","PT","CT")) +
  scale_y_continuous(labels = function(x) format(x, scientific = FALSE))


# Line graph for Real Estate and Rental and Leasing
trend_sectorField %>%
  filter(naics2_name == "Real Estate and Rental and Leasing") %>%
  ggplot(aes(x = year, y = job_count, color = DETAIL)) +
  geom_line() + geom_point(size = 5) +
  scale_x_continuous(breaks=c(2010,2011,2012,2013,2014,2015,2016,2017,2018,2019,2020)) +
  theme_gray(base_size = 15) +
  labs(title = "Trend in AI Labor Demand by AI Workforce Category (Real Estate and Rental and Leasing)",
       x = "Year",
       y = "Job Count") +
  scale_color_discrete("AI Workforce Category", breaks = c("C1","C2","PT","CT"))

# With total trend
trend_sectorField_total %>%
  filter(naics2_name == "Real Estate and Rental and Leasing") %>%
  ggplot(aes(x = year, y = job_count, color = DETAIL)) +
  geom_line() + geom_point(size = 5) +
  scale_x_continuous(breaks=c(2010,2011,2012,2013,2014,2015,2016,2017,2018,2019,2020)) +
  theme_gray(base_size = 15) +
  labs(title = "Trend in AI Labor Demand by AI Workforce Category (Real Estate and Rental and Leasing)",
       x = "Year",
       y = "Job Count") +
  scale_color_discrete("AI Workforce Category", 
                       breaks = c("Total","C1","C2","PT","CT")) +
  scale_y_continuous(labels = function(x) format(x, scientific = FALSE))

# Examine the divergence of CT job trend from overall trend in Real Estate sector by looking if any particular occupation codes are causing it
trend_realestate_ct_occ <- read_csv("trend_realestate_ct_occ_20152016.csv")

trend_realestate_ct_occ <- arrange(trend_realestate_ct_occ, year)

# Calculate differences in job count between 2015 and 2016 (a negative number indicates there were less job postings for that occupation code in 2016)
trend_realestate_ct_occ %>%
  group_by(us_occ_name) %>%
  mutate(diff = job_count - lag(job_count)) %>%
  arrange(year, diff) %>%
  filter(!is.na(diff)) %>%
  pull(us_occ_name) -> diff_occ_list

# This helps secure the order of the occupation codes in the final graph
trend_realestate_ct_occ$us_occ_name <- factor(trend_realestate_ct_occ$us_occ_name,
                                              levels = diff_occ_list)

# Side-by-side bar graph
trend_realestate_ct_occ %>%
  arrange(year) %>%
  ggplot(aes(x = us_occ_name, y = job_count, fill = as.factor(year))) +
    geom_bar(stat = "identity", position = position_dodge2(reverse = TRUE)) +
    theme_gray(base_size = 15) +
    labs(title = "Change in AI Labor Demand for CT Jobs in Real Estate (2015-2016)",
         subtitle = "Ordered from largest to smallest change",
         x = "Occupation Codes",
         y = "Job Count",
         fill = "Year") +
    coord_flip()


# Line graph for Information
trend_sectorField %>%
  filter(naics2_name == "Information") %>%
  ggplot(aes(x = year, y = job_count, color = DETAIL)) +
  geom_line() + geom_point(size = 5) +
  scale_x_continuous(breaks=c(2010,2011,2012,2013,2014,2015,2016,2017,2018,2019,2020)) +
  theme_gray(base_size = 15) +
  labs(title = "Trend in AI Labor Demand by AI Workforce Category (Information)",
       x = "Year",
       y = "Job Count") +
  scale_color_discrete("AI Workforce Category", breaks = c("C1","C2","PT","CT"))

# With total trend
trend_sectorField_total %>%
  filter(naics2_name == "Information") %>%
  ggplot(aes(x = year, y = job_count, color = DETAIL)) +
  geom_line() + geom_point(size = 5) +
  scale_x_continuous(breaks=c(2010,2011,2012,2013,2014,2015,2016,2017,2018,2019,2020)) +
  theme_gray(base_size = 15) +
  labs(title = "Trend in AI Labor Demand by AI Workforce Category (Information)",
       x = "Year",
       y = "Job Count") +
  scale_color_discrete("AI Workforce Category", 
                       breaks = c("Total","C1","C2","PT","CT")) +
  scale_y_continuous(labels = function(x) format(x, scientific = FALSE))
```


How does labor market share change across sectors?
```{r}
emp_share <- read_csv("employer_share.csv")

# Multiply by 100 to get the actual percentage
emp_share$percentage <- emp_share$percentage*100

# Group by sector and order from highest share to lowest share of job postings
emp_share <- emp_share[order(emp_share$naics2_name,
                             emp_share$percentage,
                             decreasing = T),]

# For the sole purpose of visualization, only keep the job posting data of top 3 employers in each sector
emp_share %>%
  group_by(naics2_name) %>%
  mutate(rank = rank(desc(job_count))) %>%
  mutate(job_count = ifelse(rank > 3, NA, job_count)) -> emp_share

# Change the variable to a factor with levels so that the stacked bar graph can be ordered by share size
emp_share$canon_employer <- factor(emp_share$canon_employer,
                                   levels = emp_share$canon_employer)

# Stacked bar graph
ggplot(emp_share, aes(x = naics2_name, 
                      y = percentage, 
                      fill = canon_employer)) + 
  geom_bar(stat = "identity", color="black") +
  geom_text(aes(label = canon_employer), position = position_stack(vjust = 0.7)) +
  geom_text(aes(label = job_count), position = position_stack(vjust = 0.3)) +
  theme_gray(base_size = 15) +
  theme(legend.position="none") +
  labs(title = "AI Labor Demand Broken Down by Employer in Each Sector (2019)",
       subtitle = "(Employers w/ share > 1%)",
       x = "Sector",
       y = "% of Job Postings (in Each Sector)")
```


Do different sectors tend to hire in different locations?
```{r}
location_sector <- read_csv("location_sector.csv")
location_sector <- location_sector[order(location_sector$naics2_name,
                                         desc(location_sector$job_count)),]

# Calculate percentage of job postings for each state in each sector and take the top 5
location_sector %>%
  group_by(naics2_name) %>%
  mutate(sum = sum(job_count), percentage = job_count/sum*100) %>%
  group_by(naics2_name) %>%
  slice(1:5) -> location_sector

# For the sole purpose of visualization, keep the job count information of only the top state in each sector
location_sector %>%
  group_by(naics2_name) %>%
  mutate(rank = rank(desc(job_count))) %>%
  mutate(job_count = ifelse(rank > 1, NA, job_count)) -> location_sector

# Stacked bar graph
ggplot(location_sector, aes(x = naics2_name, 
                            y = percentage, 
                            fill = canon_state)) + 
  geom_bar(stat = "identity", color="black") +
  geom_text(aes(label = canon_state), position = position_stack(vjust = 0.7)) +
  geom_text(aes(label = job_count), position = position_stack(vjust = 0.3)) +
  theme_gray(base_size = 10) +
  theme(legend.position="none") +
  labs(title = "AI Labor Demand Broken Down by Location in Each Sector (2019)",
       subtitle = "(Top 5 Locations)",
       x = "Sector",
       y = "% of Job Postings (in Each Sector)")
```


Do different sectors tend to hire in different locations? (Job count / 100,000 residents)
```{r}
location_sector <- read_csv("location_sector.csv")

# State population age 18 and over, as of July 1, 2019
state_populations <- read_csv("state_populations.csv")
state_populations <- rename(state_populations, canon_state = state)

# Merge the state population data into location_sector
location_sector <- merge(location_sector, state_populations, by="canon_state")

# Calculate job count per 100,000 residents
location_sector %>%
  mutate(relative_job_count = job_count/population*100000) -> location_sector

# Order by sector and relative job count
location_sector <- location_sector[order(location_sector$naics2_name,
                                         desc(location_sector$relative_job_count)),]

# Take the top 5 states in each sector
location_sector %>%
  group_by(naics2_name) %>%
  slice(1:5) -> location_sector

# Save the dataframe to create a table
write.csv(location_sector, "location_sector_relative.csv")

# Stacked bar graph
ggplot(location_sector, aes(x = naics2_name, 
                            y = relative_job_count, 
                            fill = canon_state)) + 
  geom_bar(stat = "identity", color="black") +
  geom_text(aes(label = canon_state), position = position_stack(vjust = 0.5)) +
  theme_gray(base_size = 10) +
  theme(legend.position="none") +
  labs(title = "AI Labor Demand Broken Down by Location in Each Sector (2019)",
       subtitle = "(Top 5 Locations; Population age >18)",
       x = "Sector",
       y = "Job Postings / 100,000 residents")
```

